---
title: "HW 1"
output:
  pdf_document: default
  html_notebook: default
---
## Homework 1 - instructions

For your first homework assignment, find a time series and plot it. Submit your plot to me. I will give you feedback *only if* your series contains anomalies that cannot be handled with the methods we have learned in class. You will analyze the series in your future homework assignments.

Please modify the notebook below:

  - Replace the example series with your chosen series.
  - Code and Markdown cells that need to be adjusted are in <span style="color:green">green</span>.
  - Save the notebook file as 'HW1_[your last name here].rmd' (e.g. HW1_sivec.rmd).
  -  Name the data file 'data_[your last name here].xls' (e.g. data_sivec.xls).
 
Submission:


  -  Submit both, the notebook and the data file (e.g. HW1_sivec.rmd \& data_sivec.xls).

**Submitt via email:vasja.sivec@uni.lu by 1.Oct.2023**


*Suggested sources*: 


https://statistiques.public.lu/en.html

    Eurostat - Statistical office of the EU features numerous series on 
    economy, finance, population, industry, environment,...

https://ec.europa.eu/eurostat/web/main/data/database  

    ECB SDW - European Central Bank data data on interest rates, loans, 
    exchange rates, financial corporations,... for european countries

https://sdw.ecb.europa.eu/

    World Bank - Global development data (rGDP, population, social 
    conditions,...) for almost all the countries in the world. 

https://data.worldbank.org/

    IMF - International Monetary Fund data features data on economic 
    and financial conditions for most countries of the world

https://www.imf.org/en/Data

    FRED data Federal Reserve Bank of St. Louis - go-to source for any 
    data on socio-economic conditions in the US

https://fred.stlouisfed.org/ 

    Penn world data University of Groningen development statistics

https://www.rug.nl/ggdc/productivity/pwt/?lang=en


Feel free to explore other sources! This could be the statistical office of your own country or a time series related to your hobby. It is advisable to select a series that features a sufficient number of observations, say at least 50.

## HW1 - hand in part

### Source and description

Full link or name of the provider: <span style="color:green">https: //fred.stlouisfed.org/series/RGDPNALUA666NRUG</span>

Series code: <span style="color:green">RGDPNALUA666NRUG</span>

Date of access: <span style="color:green">Oct. 12, 2022</span>

Description: <span style="color:green">Luxembourg gross domestic product at constant prices. Not seasonally adjusted.</span>

### Characteristics

Periodicity or frequency:<span style="color:green"> Annual</span> \
Time span:<span style="color:green"> 1950-2019</span> \
Number of observations:<span style="color:green"> 70</span> 


Plot the series and corresponding correlation and autocorrelation functions. Please plot also the first differences of the series. <span style="color:green">Modify the code below</span> so that it imports and plots your series.

```{r IMPORT}
# supress warnings
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# IMPORT DATA
# Load required packages
library(readxl)
library(ggplot2)
library(forecast)
library(ggpubr)


# Set working directory
setwd('C:\\Users\\xct385\\Desktop\\DS ws\\MADS_win23\\PART_1_intro_uni_ts\\Ex_models_R')

# Import data
df <- read_excel('rgdp_2021q4.xlsx')
y  <- ts(df[,"PIB_R_SA"],start=c(sub("*Q.","",df[1,"date"]),sub(".*Q","",df[1,"date"])),frequency=4)
```

```{r FIGURE 1}
# FIGURE 1 :  level of the series
plot_y  <- ggplot(data.frame(x = time(y), y = as.numeric(y)), aes(x = x, y = y)) + 
  geom_line() + 
  labs(title = "FIG 1 - LEVEL")
plot_ac  <- ggAcf(y,lag.max = 10) + labs(title="ACF")
plot_pac <- ggPacf(y, lag.max = 10) + labs(title="PACF")
# Display plot
ggarrange(plot_y,plot_ac,plot_pac,ncol = 3)
```

```{r FIGURE 2}
# FIGURE 2 :  1st diff of the series
y_gr <- c(100*diff(log(y)))
x    <- time(y)
plot_y  <- ggplot(data.frame(x = x[2:length(x)], y = y_gr), aes(x = x, y = y)) + 
  geom_line() + 
  labs(title = "FIG 2 - DIFF")
plot_ac  <- ggAcf(y_gr,lag.max = 10) + labs(title="ACF")
plot_pac <- ggPacf(y_gr, lag.max = 10) + labs(title="PACF")
# Display plot
ggarrange(plot_y,plot_ac,plot_pac,ncol = 3)
```